Assessment of forestry road network projection methods using the “roads” package in R

### 2022-08-23 analysis/paper/paper.Rmd
fix colour scale and ordering in figs

```{r fig1, fig.cap="Revelstoke timber supply area in British Colombia, Canada. The colour shows the cost of building roads across each pixel."}
##### Metric table bar plots
library(ggplot2)
library(dplyr)
library(stringr)
library(tidyr)
library(purrr)
library(ggpubr)
library(MetBrewer)
library(here)
library(tmap)
library(sf)

@@

pal_nm <- "Egypt"

fig_widths <- c(min = 1.18, single = 3.543, mid = 5.51, two = 7.48)

tsaBoundary <- read_sf(here(data_path_raw, "tsa27_boundaries.gpkg"))
#cost surface raster layer
bc_cost_surface <- terra::rast(here(data_path_raw, "cost_surface_bc_ha.tif"))
lakeValue <- 65000

@@

```{r fig3, fig.height=3.3, fig.width=fig_widths["single"], fig.cap="Proportional difference of projected mean from observed mean road metrics using different projection methods across the entire Revelstoke TSA (“overall”) and within cutblocks (“cutblocks”). Values greater than 0 indicate the projection overestimated the amount of roads and negative values indicate an underestimation."}
# organize data for ggplot
meanTable_long <- meanTable %>%
  pivot_longer(-c(sampleType, sampleDens, areaMean),
               names_to = "metric", values_to = "response")%>%
  mutate(sampleType = ifelse(sampleType == sampleDens, NA_character_, 
                             sampleType),
         sampleDens = paste(sampleType, sampleDens) %>% 
           factor(levels = c("NA centroid",
                             "NA observed",
                             "random low sample density",
                             "regular low sample density",
                             "random high sample density",
                             "regular high sample density",
                             "NA klementQGIS"),
                  labels = c("Centroid", 
                             "Observed",
                             "Random\nlow density",
                             "Regular\nlow density",
                             "Random\nhigh density",
                             "Regular\nhigh density",
                             "Hardy QGIS")),
         metric = factor(metric,
                         levels = c("roadDensityMean", "roadPresenceMean",
                                    "distanceToRoadMean",
                                    "forestryDisturbanceMean",
                                    "roadDisturbanceMean"),
                         labels = c("Road\ndensity", "Road\npresence",
                                    "Distance\nto road",
                                    "Forestry\ndisturbance\nfootprint",
                                    "Road\ndisturbance\nfootprint")))

@@

```{r fig5, fig.height=9, fig.width=fig_widths["two"], fig.cap="Spatial accuracy"}

matchData_sum <- matchData %>% ungroup() %>%
  mutate(agreement = factor(agreement,
                            levels = c("False negative", "False positive",
                                       "Agree roaded","Agree roadless",
                                       "Pre-existing roads")),
         sampleDens = paste(sampleType, sampleDens) %>%
           str_replace("1e-06", "low density") %>%
           str_replace("1e-04", "high density") %>%
           str_replace("centroid high density", "centroid") %>%
           str_replace("klementQGIS NA", "klementQGIS") %>%
           factor(levels = c("centroid",
                             "random low density",
                             "random high density",
                             "regular low density",
                             "regular high density",
                             "klementQGIS"),
                  labels = c("Centroid", "Random\nlow density", "Random\nhigh density",
                             "Regular\nlow density",
                             "Regular\nhigh density",
                             "Hardy QGIS")),
         metric = factor(metric,
                         levels = c("roadPresence",
                                    "forestryDisturbance",
                                    "roadDisturbance"),
                         labels = c("Road presence",
                                    "Forestry disturbance\nfootprint",
                                    "Road disturbance footprint"))) %>%
  group_by(metric, sampleDens, agreement) %>%
  summarise(count = sum(count), .groups = "drop_last") %>%
  mutate(perc = count/sum(count)*100)

>>>>

```{r setup2}
##### Metric table bar plots
library(ggplot2)
library(dplyr)
library(stringr)
library(tidyr)
library(purrr)
library(ggpubr)
library(MetBrewer)
library(here)
library(tmap)
library(sf)

@@

pal_nm <- "Redon"

fig_widths <- c(min = 1.18, single = 3.543, mid = 5.51, two = 7.48)
```

```{r fig1, fig.cap="Revelstoke timber supply area in British Colombia, Canada. The colour shows the cost of building roads across each pixel."}
tsaBoundary <- read_sf(here(data_path_raw, "tsa27_boundaries.gpkg"))
#cost surface raster layer
bc_cost_surface <- terra::rast(here(data_path_raw, "cost_surface_bc_ha.tif"))
lakeValue <- 65000

@@

```{r fig3, fig.height=4, fig.width=fig_widths["mid"], fig.cap="Proportional difference of projected mean from observed mean road metrics using different projection methods across the entire Revelstoke TSA (“overall”) and within cutblocks (“cutblocks”). Values greater than 0 indicate the projection overestimated the amount of roads and negative values indicate an underestimation."}
# organize data for ggplot
meanTable_long <- meanTable %>%
  pivot_longer(-c(sampleType, sampleDens, areaMean),
               names_to = "metric", values_to = "response")%>%
  mutate(sampleType = ifelse(sampleType == sampleDens, NA_character_, 
                             sampleType),
         sampleDens = paste(sampleType, sampleDens) %>% 
           factor(levels = c("NA centroid",
                             "NA observed",
                             "random low sample density",
                             "regular low sample density",
                             "random high sample density",
                             "regular high sample density",
                             "NA klementQGIS"),
                  labels = c("Centroid", 
                             "Observed",
                             "Random\nlow density",
                             "Regular\nlow density",
                             "Random\nhigh density",
                             "Regular\nhigh density",
                             "Hardy QGIS")),
         metric = factor(metric,
                         levels = c("roadDensityMean", "roadPresenceMean",
                                    "distanceToRoadMean",
                                    "forestryDisturbanceMean",
                                    "roadDisturbanceMean"),
                         labels = c("Road\ndensity", "Road\npresence",
                                    "Distance\nto road",
                                    "Forestry\ndisturbance\nfootprint",
                                    "Road\ndisturbance\nfootprint")))

@@

```{r fig5, fig.height=9, fig.width=fig_widths["two"], fig.cap="Comparison of performance metrics for each projection method. F measure is the harmonic mean of precision and sensitivity, precision is the proportion of  predicted roads that are observed roads and sensitivity is the proportion of observed roads that were correctly predicted."}

matchData_sum <- matchData %>% ungroup() %>%
  mutate(agreement = factor(agreement,
                            levels = c("False negative", "False positive",
                                       "Agree roaded","Agree roadless",
                                       "Pre-existing roads")),
         sampleDens = paste(sampleType, sampleDens) %>%
           str_replace("1e-06", "low density") %>%
           str_replace("1e-04", "high density") %>%
           str_replace("centroid high density", "centroid") %>%
           str_replace("klementQGIS NA", "klementQGIS") %>%
           factor(levels = c("centroid",
                             "random low density",
                             "regular low density",
                             "random high density",
                             "regular high density",
                             "klementQGIS"),
                  labels = c("Centroid", 
                             "Random\nlow density",
                             "Regular\nlow density",
                             "Random\nhigh density",
                             "Regular\nhigh density",
                             "Hardy QGIS")),
         metric = factor(metric,
                         levels = c("roadPresence",
                                    "forestryDisturbance",
                                    "roadDisturbance"),
                         labels = c("Road presence",
                                    "Forestry disturbance\nfootprint",
                                    "Road disturbance footprint"))) %>%
  group_by(metric, sampleDens, agreement) %>%
  summarise(count = sum(count), .groups = "drop_last") %>%
  mutate(perc = count/sum(count)*100)
